---
title: "Using Ecuador to Show My Mum the Effectiveness Of Vaccines"
description: |
  Watch me comb through data full of missing and nonsensical values to try and show my mother that lockdowns and vaccines are not pointless. Probably.
author:
  - name: Harriet Mason
date: 09-08-2021
output:
  distill::distill_article:
    self_contained: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.dim = c(8, 8))
#libraries
library(tidyverse)
library(lubridate)
library(RColorBrewer)
library(gt)
#devtools::install_github("RamiKrispin/coronavirus")
library(coronavirus)
#devtools::install_github("joachim-gassen/tidycovid19")
library(tidycovid19)
```

# The Anti-Vaccine Rhetoric
Unsurprisingly, my mum has been really anti-vaccine with all the Covid-19 stuff. In an effort to show that I am willing to listen to her perspective, I have watched 30 minutes of anti-vax videos over the past week and I would be lying if I said it was enjoyable. One of them was a 10 minute rant by some random guy in a Barbados town hall meeting who was talking about how Covid-19 was a man-made bio-weapon by the Chinese government. Ecuador does not have very strong data, but even crunching their numbers (there are a lot of NA values) I think I can disprove a few of the claims in the anti-vax dialogue. Ultimately a lot of the skepticism rhetoric centres around the view that the lockdowns are exessive and the vaccines are ineffective, and so, using the Ecuadorian data, we are going to check these two claims.
  
# Effectiveness of Lockdowns 

#### The Data
To assess the effectiveness of the lock downs we need a few factors. Some measure of the strictness of the lockdown, a measure of whether or not these restrictions were followed, and a measure of the number of daily cases. 

-  Check if restrictions are being followed with **Google Mobility Data**. The variables "residential" and "non-residential" are measures of the percentage change in mobility (as reported by google) when compared to the baseline period of  Jan 3 – Feb 6, 2020. Where mobility is a measure of the number of people visiting the locations in that category and the duration of these visits. The "non-residential" variable is an average of the parks, retail, groceries, transit, and related mobility, "residential" is just for people spending time in residential buildings (such as their home). Since people spent most of their time at home prior to the pandemic, the "residential" and "non-residential" variables are not directly comparable.

- Measure of lockdown strictness with **University of Oxford Government Response Tracker Data**. This gives us the stringency_index variable. This is a numeric score that summarises the strictness of closures and lockdowns. 

- Check case reports with **Johns Hopkins University Case Data**. This gives us the "cases" variable which is simply a summary of the reported cases. There is an issue in the Ecuador reporting where there are 4 days which account for a change in -9000 cases. It is impossible for there to be negative cases and so this is likely a correction for earlier mistakes in the data. Rather than average them over, there is a *large* unexplained spike very early in the dataset, I assumed this negative value was a correction for that and so I cancelled them out.


#### The Analysis

Since the point of lockdowns is to reduce the number of cases, we are going to see how the lockdown intervention measures reduced the mobility and in turn, the number of cases in Ecuador.  

```{r, table1}
# cases data
cases <- coronavirus %>%
  filter(country=="Ecuador") %>%
  filter(type=="confirmed") %>%
  mutate(cases=ifelse(cases>=0,cases,0))%>%
  select(-c(type,province, country, lat, long))
#correct for negative case thing
cases[94,2] <- cases[94,2]-(7803+1583+1480)

# google mobility
gmobility <- download_google_cmr_data(cached=TRUE) %>%
  filter(iso3c=="ECU") %>%
  mutate(nonresidential = 
           (retail_recreation + grocery_pharmacy + parks + transit_stations + workplaces)/5) %>%
  select(c(date, residential, nonresidential)) 

# lockdown indicators stuff
NPI_oxford <- download_oxford_npi_data(type="index", cached=TRUE) %>%
  filter(iso3c=="ECU") %>%
  select(c(date, stringency_index))

# make table data
lockdown_table <- cases %>%
  full_join(gmobility, by="date") %>%
  full_join(NPI_oxford, by="date")

# keep dates with values for all variables
lockdown_table <- lockdown_table[25:590,]

#plot data
plot1data <- lockdown_table %>%
  pivot_longer(cols=cases:stringency_index, names_to = "measure", values_to = "value")

#plot restrictions and stuff
#palet
mypalette<-brewer.pal(5,"YlOrRd")[2:5]
#plot
ggplot(plot1data, aes(x=date, y=value, colour=measure))+
  geom_line()+
  facet_wrap(facets=measure~., scales = "free", ncol=1) +
  theme_minimal() + 
  scale_color_manual(values=mypalette)

#table of data
table1data <- lockdown_table %>% 
  mutate(date = format(date, format = "%Y-%m")) %>%
  group_by(date) %>%
  summarise(new_cases= sum(cases),
            residential_mobility = mean(residential),
            nonresidential_mobility = mean(nonresidential),
            stringency = mean(stringency_index))

#build gt table
gt(table1data) %>% 
  tab_header(title = "Monthly Summary of Lockdown Effectiveness",
             subtitle = "Comparison of Covid-19 cases, population mobility, and lockdown restrictions averaged across each month") %>% 
  cols_label(
    date = html("Date<br>(Year-Month)"),
    new_cases = html("New Confirmed Cases"),
    residential_mobility = html("Residential Mobility <br> (% increase since 2020-01)"),
    nonresidential_mobility = html("Non-residential Mobility <br> (% increas since 2020-01)"),
    stringency = html("Oxford NPI Stringency Index"))%>%
  fmt_number(columns=new_cases:stringency, decimals = 0)


```

From this plot and table, we can see that while the restrictions were effective for a short period, this did not last. At the begining of the pandemic the citizen of Ecuador followed restrictions and stayed home, which is indicated by the higher residential mobility, lower non-residential mobility. This led to a decrease in cases. Unfortunately, this data shows that if the lockdown is strict but not managable (either because people need to work due to a lack of government support or the sheer duration of the lockdown) citizens will not follow. There is a slow increase in population mobility despite the restrictions staying the same, and obviously it leads to an increase in cases. Ultimately it seems that while the lockdown was an effective short term solution, it was not viable for the population to follow in the long run.


# Vaccine Effectiveness
### The Data
I can use the same data as before, however I will also add in a dataset that contains the vaccination rate.  
- The vaccination rate will be calculated using data from the **Our World in Data** data. The data gives total vaccinations given, and while the data has a large number of NA values, its total is equal to any I could find online, so it will be taken as correct.

- Since the vaccines also lessen the severity of Covid-19 symptoms, I will also record the reported deaths form the **Johns Hopkins University Case Data**.

### The Analysis  

```{r}
#vaccine data
owid_data <- download_owid_data(silent = FALSE, cached = TRUE)%>%
  filter(iso3c=="ECU") %>%
  mutate(total_vaccinations = ifelse(is.na(total_vaccinations),0,total_vaccinations)) %>%
  select(c(date,total_vaccinations)) %>%
  mutate(percent_of_required_dose = round((100*total_vaccinations/(2*17000000)), digits=1))

# cases data
cases <- coronavirus %>%
  filter(country=="Ecuador") %>%
  filter(!type=="recovered") %>%
  #mutate(cases=ifelse(cases>=0,cases,0))%>%
  select(-c(province, country, lat, long)) %>%
  pivot_wider(id_cols=date, names_from=type, values_from = cases)
#correct for negative case thing
cases[94,2] <- cases[94,2]-(7803+1583+1480)

# combine vaccine information
vaccine_table <- cases %>%
  full_join(owid_data, by="date") 
vaccine_table <- vaccine_table[57:588,]

# table for data
table2data <- vaccine_table %>% 
  mutate(date = format(date, format = "%Y-%m")) %>%
  group_by(date) %>%
  summarise(percent_of_required_dose = max(percent_of_required_dose, na.rm = TRUE),
            new_cases= sum(confirmed),
            new_deaths = sum(death))

#build gt table
gt(table2data) %>% 
  tab_header(title = "Monthly Summary of Vaccine Effectiveness",
             subtitle = "Comparison of Covid-19 cases, % of population vaccinated, and number of deaths reported each month") %>% 
  cols_label(
    date = html("Date<br>(Year-Month)"),
    percent_of_required_dose = html("Percent of Required Doses for 100% Vaccination"),
    new_cases = html("Total New Confirmed Cases"),
    new_deaths = html("Total New Confirmed Deaths"))%>%
  fmt_number(columns=percent_of_required_dose:new_deaths, decimals = 1)

```


While it is rather simple to see the effectiveness of the lockdowns in the previous example, the unreliable reporting of Ecuador means the effects of vaccinations are not as easy to see. The cases seem to have dropped from their peak with the higher vaccination rates, but lagged reporting means that the month with the highest deaths is when the vaccination rate is at 33%, due to deaths in prior months. Even with the weak data however, the later months show a decrease in deaths and cases (ignoring the 10,000 case spike), indicating we can still see the effectiveness of vaccines.

# Concluding Remarks
Stay home and also get vaccinated, especially if you are my 60 year old mother. 

# References
Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686,
  https://doi.org/10.21105/joss.01686
  
Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made Easy with lubridate. Journal of
  Statistical Software, 40(3), 1-25. URL https://www.jstatsoft.org/v40/i03/.
  
Erich Neuwirth (2014). RColorBrewer: ColorBrewer Palettes. R package version 1.1-2.
  https://CRAN.R-project.org/package=RColorBrewer
  
Richard Iannone, Joe Cheng and Barret Schloerke (2021). gt: Easily Create Presentation-Ready
  Display Tables. R package version 0.3.1. https://CRAN.R-project.org/package=gt
  
Rami Krispin and Jarrett Byrnes (2021). coronavirus: The 2019 Novel Coronavirus COVID-19
  (2019-nCoV) Dataset. R package version 0.3.3. https://github.com/RamiKrispin/coronavirus

Joachim Gassen (2021). tidycovid19: Download, Tidy and Visualize Covid-19 Related Data. R package
  version 0.0.0.9000.


