---
title: "Blog Post 1"
description: |
  A short description of the post.
author:
  - name: Nora Jones
    url: https://example.com/norajones
date: 09-04-2021
output:
  distill::distill_article:
    self_contained: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

# Introdction
I know I've reached a low point in my life when I'm listening to a Lowfi playlist that exclusively samples Bojack Horseman quotes. If that doesnt say I'm dead inside but I have work to do I'm not sure what does. Suffering a complete ego death unfortunately does not mean I'm free of responsibility, so here I am, sitting outside in the sun before it starts to rain and writing some post about covid data or something. Whatever.

# Data Description
```{r}
#install coronavirus package dataset
devtools::install_github("RamiKrispin/coronavirus")
library(coronavirus)
help(package="coronavirus")

ecuador_corona <- coronavirus %>%
  filter(country=="Ecuador")


#install tidycovid19 package
#devtools::install_github("joachim-gassen/tidycovid19")


```


